Challenges and Solutions of Developing and Implementing a Revolutionary Novel Desktop-as-a-Service

### 2024-04-26 paper.tex
More content (e.g. in subsection 4.4)

\begin{figure}
\includegraphics[width=\textwidth]{images/DaaS_DESIGN_Architecture_v11_english.pdf}
\caption{Architecture of the novel DaaS DESIGN.} \label{figure_architecture}
\end{figure}

@@

Examples worth mentioning in this context are eyeOS and SilveOS. The project eyeOS~\cite{liu2012research}~\cite{vidyabanu2011implementation} which simulated a desktop environment using HTML, PHP, JavaScript, and MySQL. The project SilveOS~\cite{garmpis2016design} mimics the look and feel of Windows 10 in the browser using the discontinued application framework Microsoft Silverlight. Despite being open source, both solutions have never become widely used since it is impossible to import and use native Windows or Linux applications with such solutions. All required applications have to be developed as web applications. This limitation especially caused most potential users to lose interest in web desktops because many popular native applications can hardly be replaced or re-developed, and many feature-rich and well-established Software-as-a-Service offerings like Microsoft Office 365 and Google Docs already exist.

@@

For Infrastructure-as-a-Service can be said
that our solution generally offers possibilities
to use either virtual machines or docker containers.
As can be seen in several scientific publications
discussing relevant performance aspects of such systems,
the obvious advanatge in most categories
has a lightweight container solution compared to virtual machine
if a identical program was tested in both environments
~\cite{felter2015updated}
~\cite{potdar2020performance}
~\cite{seo2014performance}.

@@

As a second observation in the area of PaaS can be added
that in most studies there is no dramatic difference
if derivates of operating systems are compared to each other.
Nevertheless, there are derivates
which are more driven towards being lightweight and more performant
~\cite{boras2020performance}  %nix vs. nix win vs. win
~\cite{balen2020performance}. %win in vm
In cases where a Windows virtual machine can be replaced with a natively operating wine platform,
additional performance potentials can be leveraged
~\cite{huang2012performance}.  % win inside vm
In contrast to that if operating system of different types are compared to each other,
there are more obvious results
especially when taking into account implementations of low-level operating system functionality,
such as interrupt handling, memory allocation, driver-specific implementations
or other more subtle requirements such as boot time
or the time needed for maintenance or system updates
~\cite{sergeev2022docker} % docker, win vs nix
~\cite{sulaiman2021comparison}. % office apps win vs nix

@@

TODO: Christian (1)
TODO: Johannes (2)

Ideen und Konzepte zur Trennung von User und Admin-Ansicht

Reduktion der Komplexität der Bedienung durch Fokus auf notwendigste Interaktionsschritte

Desktop vs. einzelne Anwendung im Browser-Tab

Anpassung des Browser-Inhalts auf die Größe der Anwendung

Herausforderungen, Entwicklung, Implementierung, Lessons-learned

>>>>

\begin{figure}
\includegraphics[width=\textwidth]{images/DaaS_DESIGN_Architecture_v11_english.pdf}
\caption{Architecture of the novel DaaS DESIGN~\cite{OJCC_2023v8i1n01_Baun}.} \label{figure_architecture}
\end{figure}

@@

Examples worth mentioning in this context are eyeOS and SilveOS. The project eyeOS~\cite{liu2012research,vidyabanu2011implementation} which simulated a desktop environment using HTML, PHP, JavaScript, and MySQL. The project SilveOS~\cite{garmpis2016design} mimics the look and feel of Windows 10 in the browser using the discontinued application framework Microsoft Silverlight. Despite being open source, both solutions have never become widely used since it is impossible to import and use native Windows or Linux applications with such solutions. All required applications have to be developed as web applications. This limitation especially caused most potential users to lose interest in web desktops because many popular native applications can hardly be replaced or re-developed, and many feature-rich and well-established Software-as-a-Service offerings like Microsoft Office 365 and Google Docs already exist.

@@

For Infrastructure-as-a-Service can be said
that our solution generally offers possibilities
to use either virtual machines or docker containers.
As can be seen in several scientific publications
discussing relevant performance aspects of such systems,
the obvious advanatge in most categories
has a lightweight container solution compared to virtual machine
if a identical program was tested in both environments
~\cite{felter2015updated,potdar2020performance,seo2014performance}.

@@

As a second observation in the area of PaaS can be added
that in most studies there is no dramatic difference
if derivates of operating systems are compared to each other.
Nevertheless, there are derivates
which are more driven towards being lightweight and more performant
~\cite{balen2020performance,boras2020performance}.  %nix vs. nix win vs. win
% ~\cite{balen2020performance}. %win in vm
In cases where a Windows virtual machine can be replaced with a natively operating wine platform,
additional performance potentials can be leveraged
~\cite{huang2012performance}.  % win inside vm
In contrast to that if operating system of different types are compared to each other,
there are more obvious results
especially when taking into account implementations of low-level operating system functionality,
such as interrupt handling, memory allocation, driver-specific implementations
or other more subtle requirements such as boot time
or the time needed for maintenance or system updates
~\cite{sergeev2022docker,sulaiman2021comparison}. % docker, win vs nix
% ~\cite{sulaiman2021comparison}. % office apps win vs nix

@@

An essential feature of a DaaS platform is its strong user-friendliness, which eliminates the complexity of the different operating systems used inside virtual machines and containers, makes the underlying hardware transparent, and simplifies the necessary interactions as much as possible.

Since users already have an operating system on their clients, and the motivation for using a DaaS is the simple use of various applications, it makes no sense for most users to have additional desktops available in the browser. The focus should be on the applications. For this reason, when developing a new DaaS solution, the main goal should be to reduce the graphical output to the limits of the respective application. The concept uses a separate browser tag for each application whose content exclusively represents the application.

Developing the automatic adaptation of the output to the limits of the respective application window is challenging. Only a few VNC and RDP servers make exporting the output of only individual windows or processes possible~\cite{OJCC_2023v8i1n01_Baun}. For Windows operating systems, the proprietary built-in RDP server and the free software TightVNC offer the export of full desktops and single windows. For Linux operating system deployments, the free software projects xrdp and x11vnc enable all desired features (see Table~\ref{tab:DaaS_Services_Overview}).

\begin{table}
	\centering
	\caption{Services used by our DaaS for Remote Access to individual Linux or Windows Applications}
	\begin{tabular}{|l|c|c|c|c|}
		\hline
		\textbf{Service}    &  \textbf{Protocol}  & \textbf{Export of}           & \textbf{Operating} & \textbf{Software}  \\
		\textbf{implementation}  &                & \textbf{single Window only}  & \textbf{System}    & \textbf{License}   \\
		\hline
		TightVNC            &  VNC                & supported                    & Windows            & GPL2               \\
		\hline
		Windows RDP server  &  RDP                & supported                    & Windows            & proprietary        \\
		\hline
		x11vnc              &  VNC                & supported                    & Linux              & GPL2               \\
		\hline
		Xrdp                &  RDP                & supported                    & Linux              & Apache 2.0         \\
		\hline
	\end{tabular}
	\label{tab:DaaS_Services_Overview}
\end{table}

The automatic adaptation of browser tabs content size, is done by.... TOTO: Johannes...

TODO: Ideen und Konzepte zur Trennung von User und Admin-Ansicht

TODO: Herausforderungen, Entwicklung, Implementierung, Lessons-learned
